---
title: "Project Structuring"
author: "April Wright"
date: "2022-10-06"
output: html_document
---

```{r setup, eval = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To be completed in lieu of class 10/10/22

For your final project in this class, you will be developing an R package. For all participants in this class, this R package will contain:

+ A function to clean some data of your choice
+ A statistical function (ANOVA, LM, other of your choosing)
+ A function that plots data

For graduate students and H-Option students, you will also include two other functions of your choosing.

Over the next few weeks, we will begin work on these packages. For this week, we will do some basic set up. 

First, on GitHub, on the left hand side of the screen, click the New Repository button. This will generate a new repository. Name it yourlastname_pkg. Copy the URL of this repo. 

Log in to RStudio. Rather than reloading your workspace as normal, click on 'New Project' in the upper right-hand side of the interface. Choose New Project from Repo. This is where you will enter the URL you pasted from GitHub. It will take a moment to generate a new, clean R install. This is where you will build your R package. Change the name from Untitled project to what you would like to the same name as your repo name.

In this directory, we will do three things. First, create a scripts directory and a data directory. Upload your data to the data directory. Within the scripts directory, create three (five for grad and H-Option) .R files (File > New File > R Script). Title each one with a descriptive name. For example, if one of your three (or five) functions will be making a histogram, you might call it `Histogram.R`.

The second thing we will do is install any needed R packages. For example, you might know you will need  tidyverse. Go nuts. Install whatever.

Lastly, we'll be setting git up. Follow the below instructions. 

```{r}
install.packages("usethis")
library(usethis)
```

Next, in the terminal:

```
git config --global user.name "Your name"
git config --global user.email "your email"
```

Then, back in R:

```{r, eval=FALSE}
use_git()
```

Select the positive option. Next, we'll create our access token.

```{r, eval=FALSE}

create_github_token()
```

Put "GBIO class project" in for purpose. Set the access period to 90 days. Create the token and copy it.

Next, we give the token to Git:

```{r, eval=FALSE}
gitcreds::gitcreds_set()

```
Paste in your token.

You may need to reload the Rstudio window, but you should now be able to commit and push your data and scripts directory.







