---
title: "ProjectOne"
output: html_document
---


# Data Project One

## 100 points
## DUE Sept 15 at 11:59 PM

Below are the questions for the first data practical assignment. This project uses the "FossilAnts.csv" file, located in the data directory for the project. The point value of each question is denoted next to it. A blank cell is below each for your answer; feel free to create more blank cells as needed.

0. (5 pts) Create a directory called `projects`, and in it, a subdirectory called `project_one`. Use `download.files` to get [this file](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_one.Rmd), and [this one](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/projects/project_one/FossilAnts.csv) and save them to the `project_one` directory.

# download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_one.Rmd", destfile = "/cloud/project/projects/project_one/project_one.Rmd")
# download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/projects/project_one/FossilAnts.csv", destfile = "/cloud/project/projects/project_one/FossilAnts.csv")

1. 5 pts. Import the `tidyverse` package and load the data. The data for this part of the practical is located in the data directory. Save the data in a variable called `project_dat`. Print the data to the screen to ensure it loaded correctly.

```{r}
library(tidyverse)
project_dat <- read_csv("FossilAnts.csv")
project_dat

```

2. (5pts) Check the datatypes of each column. There is a column called reference number. This is a static identifier - it should not be changed, and is an indexer used to identify this specimen uniquely. Do we want to treat it as an integer (this is an opinion question - answers may vary).

```
No, I personally would not want to treat the reference number as an integer but rather as a string (or character) as I'd prefer my identifiers not be numerical to avoid confusing them with anything assigned a true numeric value that, for example, we intend to run calculations on or organize and compare based on those numeric values.
```

3. (5 pts) Change the datatype of the reference number column to character. Take a peek at the function as.character().

```{r}
proj_dat <- project_dat %>% 
  mutate(reference_no = as.character(reference_no)) 
class(proj_dat$reference_no) #I found that if I did not store this pipe as another name, the reference number column remained numeric
proj_dat

```

4. (5 pts) Look at your data. What are the missing data values? In particular, have a look at the `Tribe` column. In your opinion, are these intelligent missing values for the dataset? Why or why not? If not, how would you like to change them? 

```
The missing data values in the Tribe column are listed as "None". I do not believe these are intelligent missing values for the dataset as they will not be removed with the na.omit() function. I would like to change all values listed as "None" to "NA".

```

6. (5 points) In the surveys dataset, we have genus and species split between two columns. Here, we have them combined. What are the pros and cons of the way we have recorded taxa in these two data sets.

```
The pro of having a combined genus and species column is that the taxonomic descriptors are more detailed and can be categorized easily into each species, as if you group by this column it will separate into species groups. The largest con would be that it is more difficult to sort by genus, as R will see the difference in species name and separate them based off that rather than keeping all species within a genus together.
```

7. (5pts) Please look at the help page for the `separate` function.

```{r}
?separate
```


7. (10 pts) How could you separate one column into two? 

```{r}
column_separated <- separate(data = proj_dat, col = "taxon", into = c("genus", "species"), sep = "_")
```

Next, we will test a hypothesis. Your hypothesis is that there are more specimens in the 75 million years ago (mya) - 100 mya interval than the 30 mya to the present interval.

8. (5 pts) Write out the steps you would take to address this question. Will you need to split up the data? Will you need to group the data based on the values in some column? 

Going off the notion that 1 observation, or 1 row = 1 specimen, I would create two new datasets. The first will have the specimens in the present - 30 mya interval by running a filter for both the min and max ma columns that specify values should all be less than or equal to 30 mya, ensuring the entire dataset is encompassed with every specimen that falls within our defined interval. I will obtain the number of specimens, or rows, with the command nrows(). I will do the same thing in a separate dataset for the interval 75 - 100 mya by filtering for values greater than or equal to 75 from the min and max ma columns and then obtaining the total number with nrows(). I purposefully did not include na.omit() as there was still a specimen observed no matter what other information is missing. 


9. (15 pts) Perform the operations you described in (8). 

```{r}

proj_dat_p_30 <- proj_dat %>% 
  filter(min_ma <= 30) %>% 
  filter(max_ma <= 30) %>% 
  nrow() #313

proj_dat_p_30

proj_dat_75_100 <- proj_dat %>% 
  filter(min_ma >= 75) %>% 
  filter(max_ma >= 75) %>% 
  nrow() #51

proj_dat_75_100
```

10. (5 pts) Do the results of your code support the hypothesis? 

```
No, there are more specimens in the present - 30 mya interval (n = 313) than there are specimens in the 75 - 100 mya interval (n = 51).
```


11. (10 pts) Save the dataframe with the split taxon columns into a new directory called `project_one_data_output`. Save it as a csv file called "column_separated.csv" 

```{r}
column_separated

write_csv(column_separated, path = "project_one_data_output/column_separated.csv")

```

12. (10 pts) Produce a histogram of the number of specimens by minimum age. Look at the geom_histogram() help.

```{r}
ggplot(proj_dat, aes(x = min_ma)) + geom_histogram(binwidth = 10)
  

```

13. (5 pts) Change the font size on your histogram so that it can be read comfortably on your computer screen from five feet away. Save the file as "large_font.pdf".

```{r}
ggplot(proj_dat, aes(x = min_ma)) + geom_histogram(binwidth = 10) + theme(text = element_text(size = 50))
```

14. Finally, produce a histogram of counts for each subfamily.

```{r}


ggplot(proj_dat, mapping = aes(x = subfamily)) + geom_bar(stat = "count")
       

```


# Graduate Students

1. You will do the above steps with the classroom dataset.
2. Next, you will choose three data steps from above to perform with your own data. This could be grouping, mutating, cleaning NA values, visualization, or any combination of the three. 
3. Why did you choose these steps? What function does this serve for your thesis and/or other research projects? 

For the data step below, I read in my data with function read_csv and I simultaneously assigned NA values. The NA values in my dataset are set to record as "1000", given the device records numeric data and 0 is a meaningful value in many columns, however, 1000 is not. Here I defined those as NA values.
```{r}
my_data <- read_csv("class_data_v1.csv", na = "1000")
my_data

```

For the next data step, I filtered out the na's in 3 columns using the filter(!is.na..()) command, and then I used the mutate function to create 2 new columns with the sums and means of those values.
```{r}
my_data_v2 <- my_data %>% 
  filter(!is.na(Acc_X), !is.na(Acc_Y), !is.na(Acc_Z)) %>% 
  mutate(Acc_Sum = Acc_X + Acc_Y + Acc_Z) %>% 
  mutate(Acc_Mean = mean(Acc_Sum))

my_data_v2
```

The final data step I chose was to create a historgram according to the mean values of the accelerometry data. Completing these steps is necessary for the pre-processing of my raw data before I can begin analysis. I will need to define the NA values and remove them as needed given there will almost always be an NA in each given row. It is also a crucial step for me to create new "time series" or columns from calculations using the original data columns. These are the values I will be using to create a behavioral classification model and will serve as the prediction factors for various movements and behaviors of the animal wearing the device. 
```{r}
ggplot(my_data_v2, mapping = aes(x = Acc_Mean)) + geom_histogram(binwidth = 1)

#This was not very informative as no matter the binwidth size, the histogram is uniform meaning my data is likely near uniformly distributed in terms of frequencies of values. This makes sense given the measurements being taken, so I decided to plot the means a different way.

ggplot() +
  geom_line(data = my_data_v2, mapping = aes(x = Second, y = Acc_X), color = "lightgreen") +
  geom_line(data = my_data_v2, mapping = aes(x = Second, y = Acc_Y), color = "skyblue2") +
  geom_line(data = my_data_v2, mapping = aes(x = Second, y = Acc_Z), color = "goldenrod1") +
  geom_hline(yintercept = 0) 

#Here I wanted to visualize the 3 axes of accerometry data individually plotted against time. In this case I just chose to make the y axis = seconds. However, this is because the module recording ms on the device is not synched at the moment but in the future I would like to find a way to create a time column that has continuous, sequential values at accurate intervals. Here there are many values per second and so I am curious if the ones displayed are an average, first value, etc. and how I would go about smoothing this graph out. I will eventually be averaging values over the entire second so it would be valuable to know this.

#AMW: That's pretty cool. I'm curious about the Z dimension on this one.
```


