---
output:
  pdf_document: default
  html_document: default
---
Read in and store data files; define NA values
```{r}
library(tidyverse)

XYZ_A <- read_csv("XYZ_calA.csv", na = "1000")
head(XYZ_A)

XYZ_B <- read_csv("XYZ_calB.csv", na = "1000")
head(XYZ_B)

XYZ_C <- read_csv("XYZ_calC.csv", na = "1000")
head(XYZ_C)

XYZ_D <- read_csv("XYZ_calD.csv", na = "1000")
head(XYZ_D)

XYZ_E <- read_csv("XYZ_calE.csv", na = "1000")
head(XYZ_E)

XYZ_F <- read_csv("XYZ_calF.csv", na = "1000")
head(XYZ_F)

```

Merge the data frames together with rbind() function
```{r}
XYZ_CAL <- rbind(XYZ_A, XYZ_B, XYZ_C, XYZ_D, XYZ_E, XYZ_F)
print(XYZ_CAL)

#check for error by making sure number of rows match
nrow(XYZ_CAL) #12500
nrow(XYZ_A) + nrow(XYZ_B) + nrow(XYZ_C) + nrow(XYZ_D) + nrow(XYZ_E) + nrow(XYZ_F) #12500

```

Add row ID column
```{r}
XYZ_CAL <- tibble::rowid_to_column(XYZ_CAL, "ID")
head(XYZ_CAL)

```

Create new dataframe with datetime column using make_datetime() function from Lubridate package
```{r}
#library(lubridate)

XYZ_CAL1 <- XYZ_CAL %>% 
  mutate(datetime = make_datetime(Year, Month, Day, Hour, Minute, Second))
```

Create a datetime value from string using ymd_hms() function
```{r}
DT <- ymd_hms("1997-06-10 12:06:50")
```

Visually check new column for errors
```{r}
head(XYZ_CAL1$datetime)
head(XYZ_CAL1)
```

Create a row for just the times using hms package and parse_hms() function
```{r}
library(hms)

#set format
times <- parse_hms(c("00:00:00.25", "00:00:01", "00:01:30", "01:00:00"))
times

#set value to numeric
times_num <- as.numeric(times)
times_num

#format inputs
as_hms(times_num)

#add column to dataframe
XYZ_CAL1$time <- as_hms(paste(XYZ_CAL1$Hour, XYZ_CAL1$Minute, XYZ_CAL1$Second, sep = ":"))
head(XYZ_CAL1)

```

________________________
RAW DATA PLOTS
________________________

Plot raw acceleration data from all 3 Acc XYZ columns
```{r}
rawXYZaccplot <- ggplot() + 
  geom_line(data = XYZ_CAL1, mapping = aes(x = ID, y = Acc_X, color = "X")) + 
  geom_line(data = XYZ_CAL1, mapping = aes(x = ID, y = Acc_Y, color = "Y")) + 
  geom_line(data = XYZ_CAL1, mapping = aes(x = ID, y = Acc_Z, color = "Z")) + 
  geom_hline(yintercept = 0) +
  labs(title = "Raw Acceleration", x = "sample #", y = "Triaxial Acceleration (m/s^2)") +
  theme(plot.title = element_text(hjust = 0.5))
rawXYZaccplot + theme(legend.title = element_blank())
#ggsave("/cloud/project/projects/mcgreal.pkg_project/XYZ_cal/rawXYZaccplot.png")

```

Plot raw linear acceleration data from all 3 linAcc XYZ columns
```{r}
rawXYZlinaccplot <- ggplot() + 
  geom_line(data = XYZ_CAL1, mapping = aes(x = ID, y = linAcc_X, color = "X")) + 
  geom_line(data = XYZ_CAL1, mapping = aes(x = ID, y = linAcc_Y, color = "Y")) + 
  geom_line(data = XYZ_CAL1, mapping = aes(x = ID, y = linAcc_Z, color = "Z")) + 
  geom_hline(yintercept = 0) +
  labs(title = "Raw Linear Acceleration", x = "sample #", y = "Triaxial Linear Acceleration (m/s^2)") +
  theme(plot.title = element_text(hjust = 0.5))
rawXYZlinaccplot + theme(legend.title = element_blank())
#ggsave("/cloud/project/projects/mcgreal.pkg_project/XYZ_cal/rawXYZaccplot.png")

```


________________________
SMOOTHED DATA PLOTS
________________________

Take the average of X/Y/Z measurements over 1 second; create dataframe for each
```{r}
#Acceleration columns
Xacc <- XYZ_CAL1 %>% 
  select(time, Acc_X) %>% 
  group_by(time) %>% 
  summarize(Xacc_mean = mean(Acc_X)) %>%
  group_by(time)
head(Xacc)

Yacc <- XYZ_CAL1 %>% 
  select(time, Acc_Y) %>% 
  group_by(time) %>% 
  summarize(Yacc_mean = mean(Acc_Y)) %>%
  group_by(time)
head(Yacc)

Zacc <- XYZ_CAL1 %>% 
  select(time, Acc_Z) %>% 
  group_by(time) %>% 
  summarize(Zacc_mean = mean(Acc_Z)) %>%
  group_by(time)
head(Zacc)

#Linear Acceleration columns
Xlinacc <- XYZ_CAL1 %>% 
  select(time, linAcc_X) %>% 
  group_by(time) %>% 
  summarize(Xlinacc_mean = mean(linAcc_X)) %>%
  group_by(time)
head(Xlinacc)

Ylinacc <- XYZ_CAL1 %>% 
  select(time, linAcc_Y) %>% 
  group_by(time) %>% 
  summarize(Ylinacc_mean = mean(linAcc_Y)) %>%
  group_by(time)
head(Ylinacc)

Zlinacc <- XYZ_CAL1 %>% 
  select(time, linAcc_Z) %>% 
  group_by(time) %>% 
  summarize(Zlinacc_mean = mean(linAcc_Z)) %>%
  group_by(time)
head(Zlinacc)

```

Add row ID column to each new dataframe
```{r}
Xacc <- tibble::rowid_to_column(Xacc, "ID")
head(Xacc)

Yacc <- tibble::rowid_to_column(Yacc, "ID")
head(Yacc)

Zacc <- tibble::rowid_to_column(Zacc, "ID")
head(Zacc)

Xlinacc <- tibble::rowid_to_column(Xlinacc, "ID")
head(Xlinacc)

Ylinacc <- tibble::rowid_to_column(Ylinacc, "ID")
head(Ylinacc)

Zlinacc <- tibble::rowid_to_column(Zlinacc, "ID")
head(Zlinacc)

```

Plot all acc axes together
```{r}
XYZaccplot <- ggplot() + 
  geom_line(data = Xacc, mapping = aes(x = ID, y = Xacc_mean, color = "X")) + 
  geom_line(data = Yacc, mapping = aes(x = ID, y = Yacc_mean, color = "Y")) + 
  geom_line(data = Zacc, mapping = aes(x = ID, y = Zacc_mean, color = "Z")) + 
  geom_hline(yintercept = 0) +
  labs(title = "Smoothed Acceleration", x = "time (s)", y = "Acceleration (m/s^2)") +
  theme(plot.title = element_text(hjust = 0.5))
XYZaccplot + theme(legend.title = element_blank())
ggsave("/cloud/project/projects/mcgreal.pkg_project/XYZ_cal/XYZaccplot.png")

```

Plot all linear acc together
```{r}
XYZlinaccplot <- ggplot() + 
  geom_line(data = Xlinacc, mapping = aes(x = ID, y = Xlinacc_mean, color = "X")) + 
  geom_line(data = Ylinacc, mapping = aes(x = ID, y = Ylinacc_mean, color = "Y")) + 
  geom_line(data = Zlinacc, mapping = aes(x = ID, y = Zlinacc_mean, color = "Z")) + 
  geom_hline(yintercept = 0) +
  labs(title = "Smoothed Linear Acceleration", x = "time (s)", y = "Linear Acceleration (m/s^2)") +
  theme(plot.title = element_text(hjust = 0.5))
XYZlinaccplot + theme(legend.title = element_blank())
ggsave("/cloud/project/projects/mcgreal.pkg_project/XYZ_cal/XYZlinaccplot.png")

```



_______________________________________________
Inspect linear acc######
```{r}
XYZlinaccplot <- ggplot() + 
  geom_line(data = Xlinacc, mapping = aes(x = ID, y = Xlinacc_mean, color = "X")) + 
  geom_line(data = Ylinacc, mapping = aes(x = ID, y = Ylinacc_mean, color = "Y")) + 
  geom_line(data = Zlinacc, mapping = aes(x = ID, y = Zlinacc_mean, color = "Z")) + 
  geom_hline(yintercept = 0) +
  labs(title = "Linear Acceleration", x = "time (s)", y = "Triaxial Linear Acceleration (m/s^2)") +
  theme(plot.title = element_text(hjust = 0.5))
XYZlinaccplot + theme(legend.title = element_blank())
XYZlinaccplot + xlim(0, 50)

```





_______
inspect raw acceleration
_______
raw acceleration
```{r}
rawXYZaccplot + xlim(0, 1000)

XYZlinaccplot + xlim(100, 200)

```


