Class
10 October 23
Linear Models and Stats pt 2
__________________________________

MODEL SUMMARY INTERPRETATION
__________________________________
ANova: categorical predictor
- expect sex to predict body depth
- Df: 
  the number of things that can be independent variables in your      analysis.
  always have to have at least 1 Df (total # variables -1). 
  usually its the # of variables - # parameters in model - 1. N-k-1
- Residuals: 
  red = 200 means 200 differences bw where is point and     where    is line
- Sum of the square error: 
  take every res and add it up. based on sex add it up (predicted    value) subtract expected from observed. square it to make all -    #s +
- Mean sq error: 
  sum of sq / total # D. tells you per point how much error there    is. on average how far each pt is from line. want to be low
- p value: 
  what is prob that we'd observe this distribution if there is no r   elationship between sex and body depth. < 0.05 = significant
- Adjusted r sq: has a couple mechanisms to avoid overfitting.         typically reported in papers  
  
```{r}
  
model_fit <- lm(carapace_width ~ carapace_length, data = crabs)
model_fit
summary(model_fit)

```
_____________________________
LM with multiple predictors
_____________________________
2 predictor variables. (carapace length and color) order doesn't matter just have to be after the ~
```{r}
model_fit <- lm(body_depth ~ carapace_length + color, data = crabs)
#assuming carapce length and color are independent of one another

summary(model_fit)

library(broom)
augment(model_fit) -> augmented_fit
qqnorm(augmented_fit$.resid)
qqline(augmented_fit$.resid, col = "red")

```

visualizing LM with multiple predictors
```{r}
ggplot(crabs, aes(x = carapace_length, y = body_depth, color = color)) +
  geom_point() +
  scale_color_manual(values = c("blue", "orange")) + 
  geom_smooth(method = "lm") #add lines to it showing our linear model

```

_______________________________________
LM with multiple numeric predictors
_______________________________________

```{r}
model_fit <- lm(body_depth ~ carapace_width + carapace_length, data = crabs)

summary(model_fit)
#if your crab has both carapace width and length = 0, your body depth will be 0.5
#when we increase c width by 1mm, we will decrease our body depth by almost 1/2mm (est column) assuming we have not changed c length

ggplot(crabs, aes(x = carapace_length, y = body_depth, color = carapace_width)) + 
  geom_point(size = 2.5) +
  scale_color_distiller(palette = "Reds")

```
Additive effects (+)
some variation coming from this some coming from that
____________________________________
INTERACTION EFFECTS MODEL
____________________________________

```{r}
model_fit <- lm(body_depth ~ carapace_length * color, data = crabs)
#we think bd depends on both c length and color, and that c length and color might depend on one another = interaction effect
#bd is predicted by 2 variables (c length and color) and they are not ind of one another

summary(model_fit)

#carapace_length:colororange: interaction coefficient. as you increase both of these things, you increase body depth by ___. this one is significant = we know that they are probably related to each other, there is a true relationship there. if you have a sig int coeff you need to keep it (run it as independent/additive and as interaction to see)


```

_____________________________________________
LM WITH MULTIPLE CATEGORICAL PREDICTORS
_____________________________________________

INTERACTION PLOT

```{r}
#interaction plot

crabs %>% 
  group_by(sex, color) %>% 
  summarize(mean_body_depth = mean(body_depth)) %>%  #show mean response to predictors
  ggplot(aes(x = sex, y = mean_body_depth, color = color)) +
  geom_point() +
  geom_path(aes(group = color)) + #will connect the m and f in each color group with a line
  scale_color_manual(values = c("blue", "orange")) #order is important bc color variables in alphabetical order

#on average for both sexes the orange crab is going to be deepr than blue crab. within each group, the f and m show some different dynamics

```
MANOVAS
dependent/interacting predictor variables
```{r}
model_fit <- lm(body_depth ~ sex * color, data = crabs)

summary(model_fit)

#  1.5340 independent coefficient for sex. would expect crab to be   1.5340 deeper when controlling for color but changing sex
#sexM:colororange  -1.8420 = the interaction bw color and sex is making the crab depths about -1.8420 smaller

```
