Read in data
```{r}
library(tidyverse)
Acc_X_cal <- read_csv("Acc_X.csv", na = "1000")
```

Create df with datetime column using make_datetime() function from Lubridate
```{r}
DT_df <- Acc_X_cal %>% 
  mutate(datetime = make_datetime(Year, Month, Day, Hour, Minute, Second))
```


Create a datetime value from string using ymd_hms() function
```{r}
DT <- ymd_hms("1997-06-10 12:06:50")
```

```{r}
head(DT_df$datetime)
```


```{r}
library(hms)
```

Acc_X_cal$timestamp <- ymd_hms(paste(Acc_X_cal$Year, Acc_X_cal$Month, Acc_X_cal$Day, Acc_X_cal$Hour, Acc_X_cal$Minute, Acc_X_cal$Second, sep = "-"))
View(Acc_X_cal)

```{r}
times <- parse_hms(c("00:00:00.25", "00:00:01", "00:01:30", "01:00:00"))
times

times_num <- as.numeric(times)
times_num

as_hms(times_num)
```

```{r}
DT_df$time <- as_hms(paste(DT_df$Hour, DT_df$Minute, DT_df$Second, sep = ":"))

head(DT_df)
```

Group by timestamp and take average of acc data for each column

```{r}
group_X <- DT_df %>% 
  select(time, Acc_X) %>% 
  group_by(time) %>% 
  summarize(X_mean = mean(Acc_X)) %>%
  group_by(time)

group_X
```

```{r}
group_Y <- DT_df %>% 
  select(time, Acc_Y) %>% 
  group_by(time) %>% 
  summarize(Y_mean = mean(Acc_Y)) %>%
  group_by(time)

group_Y
  
```

```{r}
group_Z <- DT_df %>% 
  select(time, Acc_Z) %>% 
  group_by(time) %>% 
  summarize(Z_mean = mean(Acc_Z)) %>%
  group_by(time)

group_Z
```
Plot all grouped Acc axes together
```{r}
Acc_grouped_plot <- ggplot() + 
  geom_line(data = group_X, mapping = aes(x = time, y = X_mean, color = "X")) + 
  geom_line(data = group_Y, mapping = aes(x = time, y = Y_mean, color = "Y")) + 
  geom_line(data = group_Z, mapping = aes(x = time, y = Z_mean, color = "Z")) + 
  geom_hline(yintercept = 0) + ylim(-2, 10)
Acc_grouped_plot + theme(legend.title = element_blank())
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_grouped_plot.png")

```

```{r}
group_X_plot <- ggplot() +
  geom_line(data = group_X, mapping = aes(x = time, y = X_mean), color = "hotpink")
group_X_plot

group_Y_plot <- ggplot() +
  geom_line(data = group_Y, mapping = aes(x = time, y = Y_mean), color = "green")
group_Y_plot

group_Z_plot <- ggplot() +
  geom_line(data = group_Z, mapping = aes(x = time, y = Z_mean), color = "goldenrod1")
group_Z_plot


```




Repeat steps for linAcc
```{r}
group_linX <- DT_df %>% 
  select(time, linAcc_X) %>% 
  group_by(time) %>% 
  summarize(linX_mean = mean(linAcc_X)) %>%
  group_by(time)
group_linX

group_linY <- DT_df %>% 
  select(time, linAcc_Y) %>% 
  group_by(time) %>% 
  summarize(linY_mean = mean(linAcc_Y)) %>%
  group_by(time)
group_linY

group_linZ <- DT_df %>% 
  select(time, linAcc_Z) %>% 
  group_by(time) %>% 
  summarize(linZ_mean = mean(linAcc_Z)) %>%
  group_by(time)
group_linZ

linAcc_grouped_plot <- ggplot() + 
  geom_line(data = group_linX, mapping = aes(x = time, y = linX_mean, color = "X")) + 
  geom_line(data = group_linY, mapping = aes(x = time, y = linY_mean, color = "Y")) + 
  geom_line(data = group_linZ, mapping = aes(x = time, y = linZ_mean, color = "Z")) + 
  geom_hline(yintercept = 0) 
linAcc_grouped_plot + theme(legend.title = element_blank())
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_grouped_plot.png")
```

```{r}
group_linX_plot <- ggplot() +
  geom_line(data = group_linX, mapping = aes(x = time, y = linX_mean), color = "hotpink") +
  ylim(0.5,-0.5) + labs(y = "X") + geom_hline(yintercept = 0, size = 0.2) 
group_linX_plot

group_linY_plot <- ggplot() +
  geom_line(data = group_linY, mapping = aes(x = time, y = linY_mean), color = "green") +
   ylim(0.5,-0.5) + labs(y = "Y") + geom_hline(yintercept = 0, size = 0.2) 
group_linY_plot

group_linZ_plot <- ggplot() +
  geom_line(data = group_linZ, mapping = aes(x = time, y = linZ_mean), color = "goldenrod1") +
   ylim(0.5,-0.5) + labs(y = "Z") + geom_hline(yintercept = 0, size = 0.2) 
group_linZ_plot

#library(cowplot)
linAcc_plots <- plot_grid(group_linX_plot, group_linY_plot, group_linZ_plot, align = "v", axis = "rlbt", nrow = 3, ncol = 1)
linAcc_plots
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_plots.png")

```

```{r}
group_X_plot <- ggplot() +
  geom_line(data = group_X, mapping = aes(x = time, y = X_mean), color = "hotpink") +
  ylim(-2, 10) + labs(y = "X") + geom_hline(yintercept = 0, size = 0.2) 
group_X_plot

group_Y_plot <- ggplot() +
  geom_line(data = group_Y, mapping = aes(x = time, y = Y_mean), color = "green") +
   ylim(-2, 10) + labs(y = "Y") + geom_hline(yintercept = 0, size = 0.2) 
group_Y_plot

group_Z_plot <- ggplot() +
  geom_line(data = group_Z, mapping = aes(x = time, y = Z_mean), color = "goldenrod1") +
   ylim(-2, 10) + labs(y = "Z") + geom_hline(yintercept = 0, size = 0.2) 
group_Z_plot

#library(cowplot)
Acc_plots <- plot_grid(group_X_plot, group_Y_plot, group_Z_plot, align = "v", axis = "rlbt", nrow = 3, ncol = 1)
Acc_plots
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_plots.png")
```

```{r}

#library(cowplot)
z1_Acc_plots <- plot_grid(group_X_plot, group_Y_plot, group_Z_plot, align = "v", axis = "rlbt", nrow = 3, ncol = 1)
z1_Acc_plots
ggsave("/cloud/project/projects/mcgreal.pkg_project/z1_Acc_plots.png")
```


OLD WAY

Create ID column
```{r}
Acc_X_cal_1 <- tibble::rowid_to_column(Acc_X_cal, "ID")

```

Plot all 3 axes of Linear Acceleration for full time duration
```{r}
linAcc_cal_plot <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_X, color = "X")) + 
  geom_line(mapping = aes(x = ID, y = linAcc_Y, color = "Y")) + 
  geom_line(mapping = aes(x = ID, y = linAcc_Z, color = "Z")) + geom_hline(yintercept = 0) + labs(title = "Forward (+X) Backward (-X) Motion", x = "Sampling interval (sample #)", y = "Linear Acceleration (m/s^2)")
linAcc_cal_plot
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_plot.png")

#graph showing the croc moving in the direction of +x. Acc is still, then accelerates (+ spike), then decelerates (- spike), then is still again
linAcc_cal_plot_2 <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_X, color = "X")) + 
  geom_line(mapping = aes(x = ID, y = linAcc_Y, color = "Y")) + 
  geom_line(mapping = aes(x = ID, y = linAcc_Z, color = "Z")) + geom_hline(yintercept = 0) + xlim(600, 800) + labs(title = "Single Movement in +X Direction", x = "Sampling interval (sample #)", y = "Linear Acceleration (m/s^2)")
linAcc_cal_plot_2
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_forward.png")

#graph showing the croc moving in the direction of +x. Acc is still, then accelerates (+ spike), then decelerates (- sprike), then is still again
linAcc_cal_plot_3 <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_X, color = "X")) + 
  geom_line(mapping = aes(x = ID, y = linAcc_Y, color = "Y")) + 
  geom_line(mapping = aes(x = ID, y = linAcc_Z, color = "Z")) + geom_hline(yintercept = 0) + xlim(1300, 1700) + labs(title = "Single Movement in -X Direction", x = "Sampling interval (sample #)", y = "Linear Acceleration (m/s^2)")
linAcc_cal_plot_3
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_backward.png")
```
###lubridate group by h m s and take mean 
Now try with Acc columns
```{r}
Acc_cal_plot <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_X, color = "X")) + 
  geom_line(mapping = aes(x = ID, y = Acc_Y, color = "Y")) + 
  geom_line(mapping = aes(x = ID, y = Acc_Z, color = "Z")) + geom_hline(yintercept = 0) + labs(title = "Forward (+X) Backward (-X) Motion", x = "Sampling interval (sample #)", y = "Acceleration (m/s^2)")
Acc_cal_plot
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_plot.png")

Acc_cal_plot_2 <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_X, color = "X")) + 
  geom_line(mapping = aes(x = ID, y = Acc_Y, color = "Y")) + 
  geom_line(mapping = aes(x = ID, y = Acc_Z, color = "Z")) + geom_hline(yintercept = 0) + xlim(600, 800) + labs(title = "Single Movement in +X Direction", x = "Sampling interval (sample #)", y = "Acceleration (m/s^2)")
Acc_cal_plot_2
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_forward.png")

Acc_cal_plot_3 <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_X, color = "X")) + 
  geom_line(mapping = aes(x = ID, y = Acc_Y, color = "Y")) + 
  geom_line(mapping = aes(x = ID, y = Acc_Z, color = "Z")) + geom_hline(yintercept = 0) + xlim(1300, 1700) + labs(title = "Single Movement in -X Direction", x = "Sampling interval (sample #)", y = "Acceleration (m/s^2)") + theme(legend.title.align = none)
Acc_cal_plot_3
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_backward.png")
```

Merge plots
```{r}
#install.packages("cowplot")
library(cowplot)

Acc_linAcc_cal <- plot_grid(linAcc_cal_plot, Acc_cal_plot)
Acc_linAcc_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_linAcc_cal.png")
```

Break up the zoomed in graphs by axis
```{r}
#Linear Acc X
linAcc_X_cal <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_X), color = "red") + 
  geom_hline(yintercept = 0) + xlim(600, 800) + labs(y = "X")
linAcc_X_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_X_cal.png")

#Linear Acc Y
linAcc_Y_cal <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_Y), color = "green") + 
  geom_hline(yintercept = 0) +
  xlim(600, 800) + labs(y = "Y")
linAcc_Y_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_Y_cal.png")

#Linear Acc Z
linAcc_Z_cal <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_Z), color = "blue") + 
  geom_hline(yintercept = 0) +
  xlim(600, 800) + labs(y = "Z")
linAcc_Z_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_Z_cal.png")

#combine them
linAcc_cal_sep <- plot_grid(linAcc_X_cal, linAcc_Y_cal, linAcc_Z_cal, labels = c('1', '1', '2', '2'), ncol = 1)
linAcc_cal_sep
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_cal_sep.png")
```

GRID_ARRANGE

break up zoomed in graphs by axes (Acc)
```{r}
#Acc X
Acc_X_cal <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_X), color = "red") + 
  geom_hline(yintercept = 0) + xlim(600, 800) + labs(y = "X")
Acc_X_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_X_cal.png")

#Linear Acc Y
Acc_Y_cal <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_Y), color = "green") + 
  geom_hline(yintercept = 0) +
  xlim(600, 800) + labs(y = "Y")
Acc_Y_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_Y_cal.png")

#Linear Acc Z
Acc_Z_cal <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_Z), color = "blue") + 
  geom_hline(yintercept = 0) +
  xlim(600, 800) + labs(y = "Z")
Acc_Z_cal
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_Z_cal.png")

#combine them
Acc_cal_sep <- plot_grid(Acc_X_cal, Acc_Y_cal, Acc_Z_cal, labels = c('1', '1', '2', '2'), ncol = 1)
Acc_cal_sep
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_cal_sep.png")
```
GROUP BY UNIQUE TIMESTAMPS AND PLOT THAT GROUP BY OBJECT



acceleration -X
```{r}
linAccX_cal_plot_3 <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = linAcc_X, color = "X")) +
  geom_hline(yintercept = 0) + xlim(1300, 1700) + labs(title = "Single Movement in -X Direction", x = "Sampling interval (sample #)", y = "Linear Acceleration (m/s^2)")
linAccX_cal_plot_3
ggsave("/cloud/project/projects/mcgreal.pkg_project/linAcc_backward_X.png")


AccX_cal_plot_3 <- Acc_X_cal_1 %>% 
  ggplot() + 
  geom_line(mapping = aes(x = ID, y = Acc_X, color = "X")) + 
   geom_hline(yintercept = 0) + xlim(1300, 1700) + labs(title = "Single Movement in -X Direction", x = "Sampling interval (sample #)", y = "Acceleration (m/s^2)") + theme(legend.title.align = none)
AccX_cal_plot_3
ggsave("/cloud/project/projects/mcgreal.pkg_project/Acc_backward_X.png")


```