---
title: "mcgreal_pkg_practice"
output: html_document
date: "2023-10-01"
---
Read in data
```{r}
mcgreal_pkg_practice <- read_csv("mcgreal.pkg.csv", na = "1000") #I made a folder in "projects" where I can practice writing code with my data for the time being
View(mcgreal_pkg_practice)

```

## R Markdown

Condense timestamp into one column
```{r}
library(tidyverse)
library(lubridate)

my_date <- ymd_hms("1997-06-10 12:06:50")

mcgreal_pkg_practice$timestamp <- ymd_hms(paste(mcgreal_pkg_practice$Year, mcgreal_pkg_practice$Month, mcgreal_pkg_practice$Day, mcgreal_pkg_practice$Hour, mcgreal_pkg_practice$Minute, mcgreal_pkg_practice$Second, sep = "-"))
View(mcgreal_pkg_practice)
```


#Create a function that selects the Acc columns and subtracts static acceleration to produce a dynamic acceleration columns for each axis
```{r}
dynamic_acc <- function(data, column, new_column){
  new_dacc <- data %>% 
    mutate(new_dacc = {{column}} - 9.8) %>% 
    rename({{new_column}} := new_dacc)
  return(new_dacc)
}

#run the function for X, Y, and Z Acc columns
#mcgreal_pkg_1 <- mcgreal_pkg_practice %>% 
  dynamic_acc(column = Acc_X, new_column = DBA_X) %>% 
  dynamic_acc(column = Acc_Y, new_column = DBA_Y) %>% 
  dynamic_acc(column = Acc_Z, new_column = DBA_Z)
View(mcgreal_pkg_1)
```

#Convert new dynamic acceleration columns to g where 1m/s^2 = 0.10197162129779283 g.
```{r}
convert_to_g <- function(data, column, new_column){
  convert <- data %>% 
    mutate(convert = {{column}} * 0.10197162129779283) %>% 
    rename({{new_column}} := convert)
  return(convert)
}

#run the function for the DBA X, Y, and Z axes
#mcgreal_pkg_2 <- mcgreal_pkg_1 %>% 
  convert_to_g(column = DBA_X, new_column = DBA_X_g) %>% 
  convert_to_g(column = DBA_Y, new_column = DBA_Y_g) %>% 
  convert_to_g(column = DBA_Z, new_column = DBA_Z_g)
View(mcgreal_pkg_2)
```


Create a function that selects a column, calculates the absolute value of each value in that column, and then produces a new column with those absolute values
```{r}
abs_values <- function(data, column, new_column){
  new <- data %>% 
    mutate(new = abs({{column}})) %>% 
    rename({{new_column}} := new) 
  return(new)
}

#run the function for X, Y, and Z Acc columns
mcgreal_pkg_practice_1 <- mcgreal_pkg_practice %>% 
  abs_values(column = Acc_X, new_column = Acc_X_abs) %>% 
  abs_values(column = Acc_Y, new_column = Acc_Y_abs) %>% 
  abs_values(column = Acc_X, new_column = Acc_Z_abs)

View(mcgreal_pkg_practice_1)
```

Create a function that adds the absolute values of 3 columns together to produce an output column of ODBA
```{r}
ODBA <- function(data, column_one, column_two, column_three, new_column){
  new <- data %>% 
    mutate(new = {{column_one}} + {{column_two}} + {{column_three}}) %>% 
    rename({{new_column}} := new)
  return(new)
}

mcgreal_pkg_practice_2 <- ODBA(data = mcgreal_pkg_practice_1, column_one = Acc_X_abs, column_two = Acc_Y_abs, column_three = Acc_Z_abs, new_column = ODBA)
View(mcgreal_pkg_practice_2)
```

Add row ID 
```{r}
mcgreal_pkg_practice_2 <- tibble::rowid_to_column(mcgreal_pkg_practice_2, "ID")
View(mcgreal_pkg_practice_2)
```

Plot Acc axes values and ODBA
```{r}
library(ggplot2)

Acc_plot <- function(df, x, yX, yY, yZ, yODBA){
  plot_XYZ <- ggplot() + 
    geom_line(data = df, mapping = aes(x = {{x}}, y = {{yX}}), color = "hotpink") + 
    geom_line(data = df, mapping = aes(x = {{x}}, y = {{yY}}), color = "skyblue2") + 
    geom_line(data = df, mapping = aes(x = {{x}}, y = {{yZ}}), color = "goldenrod1") +
    geom_line(data = df, mapping = aes(x = {{x}}, y = {{yODBA}}), color = "lightgrey") +
    geom_hline(yintercept = 0) 
  ggsave("/cloud/project/projects/mcgreal.pkg_project/plot_1.png")
  return(plot_XYZ)
}

plot_1 <- Acc_plot(df =  mcgreal_pkg_practice_2, x = ID, yX = Acc_X, yY = Acc_Y, yZ = Acc_Z ,yODBA = ODBA)
plot_1

#plot_1_1 <- mcgreal_pkg_practice_2 %>% 
  select(Acc_X, Acc_Y, Acc_Z, ODBA, timestamp) %>% 
  group_by(timestamp) %>%
  mutate(n_timestamp = n_distinct(timestamp))
#View(plot_1_1)
```

#AMW: I really love that visual.